---
output:
  word_document:
    reference_docx: template.docx
---

```{r libraries, include=FALSE, warning=FALSE, message=FALSE}
library(haven)
library(sjlabelled)
library(survival); library(survminer)
library(lme4); library(lmerTest)
library(broom.mixed)
library(gt); library(gtsummary)
library(patchwork)
library(flextable)
library(officer)
library(tidyverse)
library(ggeffects); library(ggpubr)
```

\pagebreak

1. Import demographic and laboratory data as dataframes:

```{r dm_data_import, warning=FALSE}
url <- "https://raw.githubusercontent.com/RConsortium/submissions-pilot3-adam-to-fda/main/m5/datasets/rconsortiumpilot3/tabulations/sdtm/"

# Read the DM data from public github
dm_url  <- glue::glue(url, "dm.xpt")
dm_data <- haven::read_xpt(readr::read_file_raw(dm_url))

dm_data <- dm_data %>% 
  dplyr::mutate(COUNTRY = factor(COUNTRY, levels = c("USA", "NOT USA"))) %>%
  dplyr::arrange(USUBJID)


# Read the LB data from public github
lb_url  <- glue::glue(url, "lb.xpt")
lb_data <- haven::read_xpt(readr::read_file_raw(lb_url))

lb_data   <- lb_data %>% 
  arrange(USUBJID, LBSEQ)

rm(url, dm_url, lb_url)

if (is.data.frame(dm_data)) {
  message("The data is a data frame.")
} else {
  message("The data is not a data frame.")
}
```

```{r dm_disp_header, echo=FALSE, warning=FALSE}
head(dm_data)
```

```{r lab_data_import, warning=FALSE}
if (is.data.frame(lb_data)) {
  message("The data is a data frame.")
} else {
  message("The data is not a data frame.")
}
```

```{r lab_disp_header, echo=FALSE, warning=FALSE}
head(lb_data)
```

2. Data integrity checks

2.1. Validate that all 25 variables and 306 rows are intact for demographic data:

```{r dm_check_integrity, warning=FALSE}
if (nrow(dm_data) == 306 & ncol(dm_data) == 25) {
  message("OK: The dataset has 306 rows and 25 columns, as expected.")
} else {
  message("ERROR: The dataset does not have the expected number of rows or columns.")
}
```

2.2. Validate that all 23 variables and 59580 rows are intact for laboratory data:

```{r lab_check_integrity, warning=FALSE}
if (nrow(lb_data) == 59580 & ncol(lb_data) == 23) {
  message("OK: The dataset has 59580 rows and 23 columns, as expected.")
} else {
  message("ERROR: The dataset does not have the expected number of rows or columns.")
}
```

2.3. Validate that the join between demographic and laboratory data is successful:

```{r join_data, warning=FALSE}
# Merge demographic and laboratory data
data <- lb_data %>% select(-DOMAIN) %>%
  left_join(dm_data %>% select(-DOMAIN, -SUBJID), by = c("USUBJID", "STUDYID")) %>%
  dplyr::mutate(ARM = factor(ARM)) %>%
  dplyr::select(!starts_with("LB"), starts_with("LB"))

# check if join is correct
if (nrow(data) == nrow(lb_data) & ncol(data) == ncol(dm_data) + ncol(lb_data) - 5) {
  message("The join was successful.")
} else {
  message("The join was unsuccessful.")
}
```

3. Data cleaning and summaries

3.1. Data cleaning
```{r data_cleaning_filtering}

# Define allowed visit levels
visits <- unique(data$VISIT)
ordered_visits <- visits[grepl("SCREENING|WEEK", visits)]

# Store original labels before modification
labels <- lapply(data, attr, "label")

# Data filtering and cleaning
data_clean <- data %>%
  select(USUBJID, ARM, VISIT, AGE, SEX, LBORRES, LBTEST) %>%
  filter(LBTEST == "Albumin", VISIT %in% ordered_visits) %>%
  mutate(
    LBORRES = as.numeric(LBORRES),
    VISIT   = factor(VISIT, levels = unique(c(ordered_visits, VISIT))),
    ARM     = factor(ARM, levels = unique(ARM))
  )

# Restore labels
for (col in names(labels)) {
    if (!is.null(labels[[col]]) && col %in% names(data_clean)) {
      attr(data_clean[[col]], "label") <- labels[[col]]
    }
  }

```

3.2. Baseline summary table is generated correctly by arm:

```{r baseline_summary, warning=FALSE}

baseline_summary <- data %>%
  dplyr::distinct(USUBJID, ARM, SEX, RACE, ETHNIC, COUNTRY) %>% 
  dplyr::mutate(RACE = fct_infreq(RACE),
         ETHNIC = fct_infreq(ETHNIC)) %>%
  gtsummary::tbl_summary(
    by = ARM,
    include = -USUBJID,
    digits = list(
      all_continuous() ~ c(1, 1, 1),
      all_categorical() ~ c(0,1 )),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) %>%
  gtsummary::add_p(pvalue_fun = label_style_pvalue(digits = 3)) %>%
  gtsummary::add_overall() %>%
  gtsummary::modify_header(label = "**Variable**") %>%
  gtsummary::bold_labels() %>% 
  gtsummary::as_flex_table() %>%
  flextable::theme_vanilla() %>% 
  flextable::set_table_properties(width = 1.0, layout = "autofit") %>%
  flextable::align(align = "left", part = "body") %>% 
  flextable::align(align = "left", part = "header") 

baseline_summary
```

3.3. Summaries by ARM and VISIT:

```{r summaries_by_ARM_and_VISIT, warning=FALSE}

ALB_summary <- data_clean %>% 
  dplyr::group_by(ARM, VISIT) %>%
  dplyr::summarise(n      = n(),
                   mean   = round(mean(LBORRES, na.rm = TRUE), 2),
                   sd     = round(sd(LBORRES, na.rm = TRUE), 2),
                   median = median(LBORRES, na.rm = TRUE),
                   min    = min(LBORRES, na.rm = TRUE),
                   max    = max(LBORRES, na.rm = TRUE),
                   .groups = "drop") %>%
  flextable::flextable() %>%
  flextable::theme_booktabs() %>%
  flextable::autofit()

ALB_summary
```

4. Statistical modeling

4.1 LM model
```{r lm_model, warning=FALSE}

# LM model ----
lm_model <- lm(LBORRES ~ ARM * VISIT + AGE, 
                data = data_clean)

lm_model_results  <- broom.mixed::tidy(lm_model) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3)))

lm_numeric_cols <- names(lm_model_results)[sapply(lm_model_results, is.numeric)]

lm_model_table <- lm_model_results %>%
  flextable() %>%
  colformat_num(j = lm_numeric_cols, digits = 3) %>%  # Ensure 3-digit precision
  theme_booktabs() %>%  # Optional: Clean table style
  autofit()  # Adjust column widths
lm_model_table
```

4.2 LMM model

```{r llm_model, warning=FALSE}

# LMM model ----
# fixed effects: ARM, VISIT, AGE
# random effect: USUBJID
lmm_model <- lmer(LBORRES ~ ARM * VISIT + AGE + (1 | USUBJID), 
              data = data_clean)

lmm_model_results  <- broom.mixed::tidy(lmm_model) %>%
  mutate(across(where(is.numeric), ~ round(.x, 1))) %>% 
  select(-group) %>% 
  filter(term %in% c("VISITWEEK 2", "VISITWEEK 4", "	VISITWEEK 6", "VISITWEEK 8", "VISITWEEK 12", "VISITWEEK 16", "VISITWEEK 20", "VISITWEEK 24", "VISITWEEK 26", "AGE"))

lmm_numeric_cols <- names(lmm_model_results)[sapply(lmm_model_results, is.numeric)]

lmm_model_table <- lmm_model_results %>%
  flextable() %>%
  colformat_num(j = lmm_numeric_cols, digits = 1) %>%  # Ensure 3-digit precision
  theme_booktabs() %>%  # Optional: Clean table style
  autofit()  # Adjust column widths
lmm_model_table
```

4.3. LMM model with nesting and map function

```{r}
data_perf <- data %>%
  select(USUBJID, ARM, VISIT, AGE, LBORRES, LBTEST) %>%
  filter(VISIT %in% ordered_visits) %>%
  mutate(
    LBORRES = as.numeric(LBORRES),
    VISIT   = factor(VISIT, levels = unique(c(ordered_visits, VISIT))),
    ARM     = factor(ARM, levels = unique(ARM))
  ) %>%
  filter(!is.na(LBORRES))
 
tests <- unique(data_perf$LBTEST)[1:5]
 
data_perf_filtered <- data_perf %>% filter(LBTEST %in% tests)
 
# Analysis with nesting and map function (replacement for for loop)
data_analysis <- data_perf_filtered %>%
  group_by(LBTEST) %>%
  nest() %>%
  mutate(
    model        = map(.x = data, ~ lmer(LBORRES ~ ARM * VISIT + AGE + (1 | USUBJID), data = .)),  # Fit the model once
    test_results = map(model, tidy),  # Extract tidy results
    tables       = map(model, ~ tbl_regression(., exponentiate = FALSE) # Generate regression table
                      )
    )
# unnest the 'test_results'
test_results <- data_analysis %>%
  select(LBTEST, test_results) %>%
  unnest(test_results)
test_results
 
```



5. Laboratory results by ARM and VISIT and predicted effects of ARM

```{r boxplot, warning=FALSE, fig.width=6.5, fig.height=4.5, fig.cap="Combined plot of laboratory results by arm and visit and predicted effects of arm on laboratory results."}

plot1 <- data_clean %>%
  filter(VISIT %in% c("SCREENING 1", "WEEK 12", "WEEK 26")) %>% 
  ggplot(aes(x = VISIT, y = LBORRES, fill = ARM)) +
  geom_boxplot() +
  labs(
    title = "Laboratory Results by Arm and Visit",
    x = "Visit",
    y = "Laboratory Results"
    ) +
  theme_minimal() +
  coord_cartesian()

interaction_effects <- ggeffects::ggpredict(lmm_model, terms = c("ARM"))

plot2 <- interaction_effects %>%
  plot() +
  labs(
    title = "Predicted Effects: Arm x Laboratory Results",
    x = "Arm",
    y = "Laboratory Results"
  ) +
  theme_minimal()

combined_plot <- plot1 + plot2 + plot_layout(ncol = 1)
combined_plot
```


6. Measure system response time for queries:

```{r performance, warning=FALSE}

data_perf <- data %>%
  select(USUBJID, ARM, VISIT, AGE, LBORRES, LBTEST) %>%
  filter(VISIT %in% ordered_visits) %>%
  mutate(
    LBORRES = as.numeric(LBORRES),
    VISIT   = factor(VISIT, levels = unique(c(ordered_visits, VISIT))),
    ARM     = factor(ARM, levels = unique(ARM))
  ) %>%
  filter(!is.na(LBORRES))

tests <- unique(data_perf$LBTEST)[1:30]

start_time <- Sys.time()

data_perf_filtered <- data_perf %>% filter(LBTEST %in% tests)

# Analysis with nesting and map function (replacement for for loop)
data_analysis <- data_perf_filtered %>%
  group_by(LBTEST) %>%
  nest() %>%
  mutate(test_results = map(.x=data, ~ lmer(LBORRES ~ ARM * VISIT + AGE + (1 | USUBJID),
                                            data = .) %>% tidy()
                            )
         ) %>%
  unnest(test_results) %>%
  select(-data)


for (i in tests) {
  lmer(LBORRES ~ ARM * VISIT + AGE + (1 | USUBJID),
     data = data_perf %>% filter(LBTEST == i)) %>% na.omit()
  #print(i)
  }
end_time <- Sys.time()
response_time <- end_time - start_time
print(paste("Total Time:", response_time))
```

