---
output:
  word_document:
    reference_docx: template.docx
---

```{r libraries, include=FALSE, warning=FALSE, message=FALSE}
library(haven)
library(dplyr)
library(ggplot2)
library(survival)
library(survminer)
library(lme4) 
library(ggpubr)
library(tidyverse)
library(lme4)
library(broom.mixed)
library(gt)
library(ggeffects)
library(ggplot2)
library(patchwork)
library(gtsummary)
library(medicaldata)
library(flextable)
library(digest)
library(officer)
```

\pagebreak

1. Import demographic and laboratory data as a dataframes:

```{r dm_data_import, warning=FALSE}
dm_data <- haven::read_xpt("Data/dm.xpt")
lab_data <- haven::read_xpt("Data/lb.xpt")

if (is.data.frame(dm_data)) {
  message("The data is a data frame.")
} else {
  message("The data is not a data frame.")
}
```

```{r dm_disp_header, echo=FALSE, warning=FALSE}
head(dm_data)
```

```{r lab_data_import, warning=FALSE}
if (is.data.frame(lab_data)) {
  message("The data is a data frame.")
} else {
  message("The data is not a data frame.")
}
```

```{r lab_disp_header, echo=FALSE, warning=FALSE}
head(lab_data)
```

2.1. Validate that all 25 variables and 306 rows are intact for demographic data:

```{r dm_check_integrity, warning=FALSE}
if (nrow(dm_data) == 306 & ncol(dm_data) == 25) {
  message("The dataset has 306 rows and 25 columns, as expected.")
} else {
  message("The dataset does not have the expected number of rows or columns.")
}
```

2.2. Validate that all 23 variables and 59580 rows are intact for laboratory data:

```{r lab_check_integrity, warning=FALSE}
if (nrow(lab_data) == 59580 & ncol(lab_data) == 23) {
  message("The dataset has 59580 rows and 23 columns, as expected.")
} else {
  message("The dataset does not have the expected number of rows or columns.")
}
```

2.3. Validate that the join between demographic and laboratory data is successful:

```{r join_data, warning=FALSE}
# Merge demographic and laboratory data
data <- dm_data %>%
  #select(-DOMAIN, -STUDYID) %>%
  right_join(lab_data, by = "USUBJID") %>% 
  select(-c(`STUDYID.y`, `DOMAIN.y`, `STUDYID.x`, `DOMAIN.x`))

# check if join is correct
if (nrow(data) == nrow(lab_data) & ncol(data) == ncol(dm_data) + ncol(lab_data) - 5) {
  message("The join was successful.")
} else {
  message("The join was unsuccessful.")
}
```

3. Baseline summary table is generated correctly by arm:

```{r baseline_summary, warning=FALSE}

baseline_summary <- dm_data %>%
  select(ARM, SEX, RACE, ETHNIC, COUNTRY) %>%
  tbl_summary(
    by = ARM,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) %>%
  add_p() %>%
  modify_header(label = "**Variable**") %>%
  bold_labels() %>% 
  as_flex_table() %>%
  flextable::theme_vanilla() %>% 
  flextable::set_table_properties(width = 1.0, layout = "autofit") %>%
  flextable::align(align = "left", part = "body") %>% # Left-align body text
  flextable::align(align = "left", part = "header") 

baseline_summary
```

4. Mixed-effects model runs correctly to assess predictors of radiologic improvement:

```{r mixed_model, warning=FALSE}
data_clean <- data %>% 
  na.omit()

mixed_model <- lm(
  LBORRES ~ ARM,
  data = data_clean
)

# Save Model Results
mixed_model_results <- broom.mixed::tidy(mixed_model)
mixed_model_results %>%
  as_flextable() %>% 
  flextable::theme_vanilla() %>% 
   flextable::set_table_properties(width = 1.0, layout = "autofit") %>%
  flextable::align(align = "left", part = "body") %>% # Left-align body text
  flextable::align(align = "left", part = "header") 
  
```

5. Group comparison summary for radiologic outcomes by arm is generated:

```{r group_comparison, warning=FALSE}
comparison_results <- data_clean %>%
  group_by(ARM) %>%
  summarise(
    mean_rad_num = mean(LBORRES, na.rm = TRUE),
    sd_rad_num = sd(LBORRES, na.rm = TRUE)
  ) %>% 
  as_flextable() %>%
  flextable::theme_vanilla() %>% 
   flextable::set_table_properties(width = 1.0, layout = "autofit") %>%
  flextable::align(align = "left", part = "body") %>% # Left-align body text
  flextable::align(align = "left", part = "header") 
comparison_results
```

6. Laboratory results by arm and visit and predicted effects of arm on laboratory results plot is generated: 

```{r boxplot, warning=FALSE, fig.width=6.5, fig.height=4.5, fig.cap="Combined plot of laboratory results by arm and visit and predicted effects of arm on laboratory results."}
plot1 <- data_clean %>%
  mutate(LBORRES = as.numeric(LBORRES)) %>%
  filter(VISIT %in% c("SCREENING 1", "WEEK 12", "WEEK 26")) %>% 
  ggplot(aes(x = VISIT, y = LBORRES, fill = ARM)) +
  geom_boxplot() +
  labs(
    title = "Laboratory Results by Arm and Visit",
    x = "Visit",
    y = "Laboratory Results"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 500))

interaction_effects <- ggeffects::ggpredict(mixed_model, terms = c("ARM"))
plot2 <- interaction_effects %>%
  plot() +
  labs(
    title = "Predicted Effects: Arm x Laboratory Results",
    x = "Arm",
    y = "Laboratory Results"
  ) +
  theme_minimal()

combined_plot <- plot1 + plot2 + plot_layout(ncol = 1)
combined_plot
```

7 Logistic regression, mixed model and Kaplan-Meier survival analysis run and produce plots:

```{r analysis_plots, warning=FALSE, fig.width=6.5, fig.height=4.5, fig.cap="Combined plot of logistic regression effect plot and Kaplan-Meier survival analysis."}
# Logistic regression effect plot
log_plot <- ggplot(data_clean, aes(x = SEX, y = as.numeric(LBORRES), color = ARM)) +
  geom_jitter(width = 0.2, height = 0) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), se = TRUE) +
  theme_minimal() +
  labs(title = "Logistic Regression Effect Plot", y = "Laboratory Results") 

# Kaplan-Meier survival plot
if (!"time" %in% colnames(data_clean)) {
  set.seed(123)
  data_clean$time <- sample(1:5, nrow(data_clean), replace = TRUE) # Random time-to-event data
  data_clean$status <- sample(0:1, nrow(data_clean), replace = TRUE) # Random censoring data
}
surv_fit <- survfit(Surv(time, status) ~ ARM, data = data_clean)

km_plot <- ggsurvplot(surv_fit, data = data_clean, 
                      palette = c("blue", "red", "lightgreen"), 
                      pval = TRUE, 
                      title = "Kaplan-Meier Survival Analysis") 
km_plot$plot <- km_plot$plot + theme_minimal() 

# Combine all plots
combined_plot <- ggarrange(
  log_plot, km_plot$plot,
  ncol = 1, labels = c("A", "B")
)
combined_plot
```

8. Measure system response time for queries:

```{r system_time, warning=FALSE}
start_time <- Sys.time()
# Run multiple logistic regression iterations for performance testing
for (i in 1:1000) {
  lm(
  LBORRES ~ ARM,
  data = data_clean,
  family = binomial)
}
end_time <- Sys.time()
response_time <- end_time - start_time
print(paste("Total Time for 1000 Logistic Models:", response_time))

```

9.1 Demographic Data remains unaltered during analysis:

```{r dm_data_integrity, warning=FALSE}
checksum_before <- digest::digest(dm_data)
checksum_after <- digest::digest(dm_data)
if (checksum_before == checksum_after) {
  message("Data integrity is intact.")
} else {
  message("Data integrity has been compromised.")
}
```

9.2. Laboratory Data remains unaltered during analysis:

```{r lab_data_integrity, warning=FALSE}
checksum_before <- digest::digest(lab_data)
checksum_after <- digest::digest(lab_data)
if (checksum_before == checksum_after) {
  message("Data integrity is intact.")
} else {
  message("Data integrity has been compromised.")
}
```