---
title: "R, RStudio, R packages Installation and Updating Instruction"
output: 
  word_document:
    reference_docx: template.docx
---

## R Installation 

1. On [R CRAN website](https://cran.r-project.org/) in "Download and Install R" section choose an appropriate link for the software, click "install R for the first time".

## RStudio Installation

1. On [RStudio website](https://www.rstudio.com/products/rstudio/download/) click "Installers RStudio" and install RStudio for Windows. For other software and installation options scroll below.

2. Open RStudio.

## R and RStudion Updating

1. Close current R session. 

2. On [R CRAN website](https://cran.r-project.org/) in "Download and Install R" section choose an appropriate link for the software, click "install R for the first time".

3. On [RStudio website](https://www.rstudio.com/products/rstudio/download/) click "Installers RStudio" and install RStudio for Windows. For other software and installation options scroll below.

4. Setup new vwrsion of RStudio.

5. Reopen R Session in RStudio and choose the newest R version in Tools - Global options General - R Session - R version - Change - Apply.

6. Close and reopen RStudio.

7. Check information about session to ensure that the newest R and RStudio versions are installed. 

- Type in the console:

``` {r sessinf, eval=FALSE, echo=TRUE}

utils::sessionInfo()

```

to check location, timezome, locale, base packages, loaded packages.

- Type in the console:

``` {r rvers, eval=FALSE, echo=TRUE}

R.Version()

```

to check R version (full name and release date), OS, platform, and other information.

- Type in the console:

``` {r versinfo, eval=FALSE, echo=TRUE}

rstudioapi::versionInfo()

```

to check RStudio version, build, and other information.

## R packages Installation

1. Install R packages from CRAN repository.

- Go to Packages - Install - Type package name - Install OR

- Type in the console:

``` {r install, eval=FALSE, echo=TRUE}

install.packages("package_name")

```

to install a new package.

2. Load R packages.

- Type in the console:

``` {r load, eval=FALSE, echo=TRUE}

library(package_name)

```

to load a package in the environment.

## R packages Updating

Go to Tools - Check for Package updates - Select all - Install updates.
