---
title: "Renv Instruction"
output:
  word_document:
    reference_docx: template.docx
---

## What is renv?

Package that helps create reproducible environments for your R projects.

More info: [Introduction to renv](https://rstudio.github.io/renv/articles/renv.html?_gl=1*1lvqqs5*_ga*MTYzNjAxNDk2MS4xNzIyNTAzNDcx*_ga_2C0WZ1JHG0*MTczNDQzNjAzMi4xMi4wLjE3MzQ0MzYwMzIuMC4wLjA.).

## Workflow with renv

1.  Install renv package:

``` {r install, eval=FALSE, echo=TRUE}

install.packages("renv")

```

2.  Initialize renv:

``` {r init, eval=FALSE, echo=TRUE}

renv::init()

```

Outcome:

-   renv creates a new folder with project-specific library for your packages - renv/library.

-   It generates a renv.lock file, which records the exact versions of all currently installed packages.

-   .Rprofile file is also created to automatically activate renv when the project is loaded.

3.  Install new packages or update existing ones:

```{r packages, eval=FALSE, echo=TRUE}

install.packages("wrappedtools")

```

4.  Snapshot environment to the renv.lock file:

```{r snap, eval=FALSE, echo=TRUE}

renv::snapshot()

```

5.  Restore environment of the project or share the project with a new person:

```{r restore, eval=FALSE, echo=TRUE}

renv::restore()

```

Outcome:

-   renv reads the renv.lock file.

-   It installs the exact versions of the required packages into the project library.

6.  To share with a new person give access to the entire project or a specific script along with renv.lock file. Use renv::restore() to recreate the exact package environment as recorded in the renv.lock file.
7.  Use renv::status() to check for issues or unused packages.
